An event-sourced and decentralized database with IPLD and materialized views

### 2019-09-23 draft.tex
gets rid of all "see Section"
Signed-off-by: andrew <andrew@textile.io>

A key principal of ES and append only logs is that all changes to application state are stored as a sequence of events. Because any given state is simply the result of a series of atomic updates, the log can be used to reconstruct past states or process retroactive updates \cite{fowlerEventSourcing}. The same principal means a log can be viewed as a mechanism to support an infinite number of valid state interpretations (see Section  \ref{sec:ViewsProjections}). In other words, with minimal conformity, a single log can model multiple application states \cite{microsoftcorporationAzureApplicationArchitecture}.

@@

Event-driven systems (e.g., event sourcing) often make use of ACLs with some distinct properties. The ACL of an event-driven system is usually a list of access rules built from a series of events. For example, the two events, ``grant Bob write access'' and ``revoke read access from Alice'' would together result in a final ACL state where, Bob has read and write access, but Alice does not (see Section  \ref{sec:ViewsProjections}).

@@

\begin{definition}
(Reader). Any Peer capable of reading a Log. Practically speaking, this means any Peer with the Log's Read Key (see Section  \ref{sec:KeysEncryption}).
\end{definition}

@@

Together with a cryptographic signature, an Event is written to a log with an additional node (see Figure \ref{fig:EventLog}) enabling log verification by readers (see Section \ref{sec:KeysEncryption}).

@@

\begin{definition} (Follower). Log Writers can designate other IPFS Peers to ``follow'' a Log, potentially replicating and/or republishing Events. A Follower is capable of receiving Log updates and traversing linkages via the Follow Key (see Section  \ref{sec:KeysEncryption}), but is not able to read the Log's contents. Followers should be server-based --- i.e., always online and behind a public IP address.\end{definition}

@@

In practice, Writers are solely responsible for announcing their Log's addresses. This ensures a conflict-free address list without additional complexity. Some Followers may be in the business of replicating Logs (see section  \ref{sec:Bots}), in which case Writers will announce the additional Log address to Readers. This allows them to \emph{pull} (or subscribe to push-based) Events from the Follower's Log address when the Writer is offline or unreachable (Figure \ref{fig:Pulling}).

@@

Textile Logs are designed to be shared, composed, and layered into datasets (Figure \ref{fig:LogEncryption}). Therefore, Logs are encrypted by default in a manner that enables access control (see Section  \ref{sec:AccessControl}) and the Follower mechanism discussed in the previous section.

@@

In Textile, the \emph{interface} to Logs is managed as a Thread. As mentioned previously, a Thread is a collection of Logs on a given topic. Threads are an event-sourced, distributed database, and can be used to maintain a single, collaboratively edited, watched, or hosted dataset across multiple Peers. Threads provide the mechanism to combine multiple Logs from individual authors into singular shared states through the use of either cross-Log sequencing (e.g. using a Bloom Clock, Merkle-Clock, or Hybrid Logical Clock \cite{kulkarniLogicalPhysicalClocks2014}) or a CRDT (see Section \ref{sec:CRDTs}).

@@

\begin{definition}
(Access-Controlled): This variant declares that consumers should assume an access control list is composable from Log Events. The ACL represents a permissions rule set that must be applied when reading data (see Section  \ref{sec:AccessControl} and Example \ref{ex:Identity}b).
\end{definition}

@@

\begin{definition}
(Invite). An Event containing a mapping of Log IDs to Key Sets, which can be used to join a Thread. Threads backed by an access control list (see Section  \ref{sec:AccessControl}) will also include the current ACL for the Thread in an Invite. This enables Peers to invite others to only read or follow a Thread, instead of becoming a full-on collaborator, i.e., a new Log Writer.
\end{definition}

@@

The notion of the Follow Key (see Section  \ref{sec:KeysEncryption}) makes duplicating all Log Events trivial. This allows any Peer in the network to be granted the responsibility of replicating data from another Peer without having read access to the data contained within the Log entries. This type of Log replication can act as a data backup mechanism. It can also be used to build services that react to Log Events, potentially pushing data to disparate, non-Textile systems, especially if the replication service is granted read access to the Log Events (see Section  \ref{sec:LogSync}).

@@

In practice, Snapshots are written to their own Event log and stored locally. They can potentially be synced (see Section \ref{sec:LogSync}) to other peers as a form of data backup or to optimize state initialization when a new peer starts participating in a shared Thread (saving disk space, bandwidth, and time). They can similarly be used for initializing state during recovery.

@@

One of the most important properties of a shared data model is the ability to apply access control rules. There are two forms of access control possible in Threads, Sequence-level ACLs and Thread-level ACLs. Thread-level access control lists (ACLs) allow creators to specify who can \emph{follow, read, write, and delete} Thread data. Similarly, sequence-level ACLs provide more granular control to Thread-writers on a per-sequence (see Definition \ref{ex:SequenceID}) basis. Both types of ACLs are materialized views, or more specifically, JSON CRDT documents, built from Log Events (see Section \ref{sec:TexCRDT}). ACLs implemented as JSON documents provides two advantages over static or external ACL rules (although static and external ACLs are also possible). First, ACLs are fully mutable, allowing developers to create advanced rules for collaboration with any combination of readers, writers, and followers. Second, because ACLs are JSON documents, ACLs can specify their own editing rules (i.e. allowing multiple Thread participants to modify the ACL) in a self-referencing way.

>>>>

A key principal of ES and append only logs is that all changes to application state are stored as a sequence of events. Because any given state is simply the result of a series of atomic updates, the log can be used to reconstruct past states or process retroactive updates \cite{fowlerEventSourcing}. The same principal means a log can be viewed as a mechanism to support an infinite number of valid state interpretations (Section  \ref{sec:ViewsProjections}). In other words, with minimal conformity, a single log can model multiple application states \cite{microsoftcorporationAzureApplicationArchitecture}.

@@

Event-driven systems (e.g., event sourcing) often make use of ACLs with some distinct properties. The ACL of an event-driven system is usually a list of access rules built from a series of events. For example, the two events, ``grant Bob write access'' and ``revoke read access from Alice'' would together result in a final ACL state where, Bob has read and write access, but Alice does not (Section  \ref{sec:ViewsProjections}).

@@

\begin{definition}
(Reader). Any Peer capable of reading a Log. Practically speaking, this means any Peer with the Log's Read Key (Section  \ref{sec:KeysEncryption}).
\end{definition}

@@

Together with a cryptographic signature, an Event is written to a log with an additional node (see Figure \ref{fig:EventLog}) enabling log verification by readers (Section \ref{sec:KeysEncryption}).

@@

\begin{definition} (Follower). Log Writers can designate other IPFS Peers to ``follow'' a Log, potentially replicating and/or republishing Events. A Follower is capable of receiving Log updates and traversing linkages via the Follow Key (Section  \ref{sec:KeysEncryption}), but is not able to read the Log's contents. Followers should be server-based --- i.e., always online and behind a public IP address.\end{definition}

@@

In practice, Writers are solely responsible for announcing their Log's addresses. This ensures a conflict-free address list without additional complexity. Some Followers may be in the business of replicating Logs (Section  \ref{sec:Bots}), in which case Writers will announce the additional Log address to Readers. This allows them to \emph{pull} (or subscribe to push-based) Events from the Follower's Log address when the Writer is offline or unreachable (Figure \ref{fig:Pulling}).

@@

Textile Logs are designed to be shared, composed, and layered into datasets (Figure \ref{fig:LogEncryption}). Therefore, Logs are encrypted by default in a manner that enables access control (Section  \ref{sec:AccessControl}) and the Follower mechanism discussed in the previous section.

@@

In Textile, the \emph{interface} to Logs is managed as a Thread. As mentioned previously, a Thread is a collection of Logs on a given topic. Threads are an event-sourced, distributed database, and can be used to maintain a single, collaboratively edited, watched, or hosted dataset across multiple Peers. Threads provide the mechanism to combine multiple Logs from individual authors into singular shared states through the use of either cross-Log sequencing (e.g. using a Bloom Clock, Merkle-Clock, or Hybrid Logical Clock \cite{kulkarniLogicalPhysicalClocks2014}) or a CRDT (Section \ref{sec:CRDTs}).

@@

\begin{definition}
(Access-Controlled): This variant declares that consumers should assume an access control list is composable from Log Events. The ACL represents a permissions rule set that must be applied when reading data (Section  \ref{sec:AccessControl} and Example \ref{ex:Identity}b).
\end{definition}

@@

\begin{definition}
(Invite). An Event containing a mapping of Log IDs to Key Sets, which can be used to join a Thread. Threads backed by an access control list (Section  \ref{sec:AccessControl}) will also include the current ACL for the Thread in an Invite. This enables Peers to invite others to only read or follow a Thread, instead of becoming a full-on collaborator, i.e., a new Log Writer.
\end{definition}

@@

The notion of the Follow Key (Section  \ref{sec:KeysEncryption}) makes duplicating all Log Events trivial. This allows any Peer in the network to be granted the responsibility of replicating data from another Peer without having read access to the data contained within the Log entries. This type of Log replication can act as a data backup mechanism. It can also be used to build services that react to Log Events, potentially pushing data to disparate, non-Textile systems, especially if the replication service is granted read access to the Log Events (Section  \ref{sec:LogSync}).

@@

In practice, Snapshots are written to their own Event log and stored locally. They can potentially be synced (Section \ref{sec:LogSync}) to other peers as a form of data backup or to optimize state initialization when a new peer starts participating in a shared Thread (saving disk space, bandwidth, and time). They can similarly be used for initializing state during recovery.

@@

One of the most important properties of a shared data model is the ability to apply access control rules. There are two forms of access control possible in Threads, Sequence-level ACLs and Thread-level ACLs. Thread-level access control lists (ACLs) allow creators to specify who can \emph{follow, read, write, and delete} Thread data. Similarly, sequence-level ACLs provide more granular control to Thread-writers on a per-sequence (see Definition \ref{ex:SequenceID}) basis. Both types of ACLs are materialized views, or more specifically, JSON CRDT documents, built from Log Events (Section \ref{sec:TexCRDT}). ACLs implemented as JSON documents provides two advantages over static or external ACL rules (although static and external ACLs are also possible). First, ACLs are fully mutable, allowing developers to create advanced rules for collaboration with any combination of readers, writers, and followers. Second, because ACLs are JSON documents, ACLs can specify their own editing rules (i.e. allowing multiple Thread participants to modify the ACL) in a self-referencing way.
